---
title: "STAT 420 — Project"
author: "Winston Chen, Daniel hooks, Naidan Ganbold, Erik Janc, Michael Lee"
date: "November 19, 2021"
output:
  html_notebook: default
  pdf_document: default
---
```{r setup, include=FALSE}
library(finalfit)
library(knitr)
library(tidyverse)
library(tidyquant)

knitr::opts_chunk$set(echo=TRUE, error=TRUE, eval=TRUE)
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```
## Introduction

## Data Preprocessing
Read the initial CSV file:
```{r}
pollution_raw = read.csv(
  "./data/pollution_us_2000_2016.csv",
  header=TRUE,
)
colnames(pollution)
```
Only include records for the following cities:

* Chicago
* Denver
* Los Angeles
* New York City

Additionally, for ease of viewing, drop irrelevant columns that we are not analyzing.
```{r}
pollution = subset(
  pollution_raw,
  pollution_raw$City %in% c("Chicago", "Denver", "Los Angeles", "New York City"),
  drop=TRUE
)

#cols = c("State", "County", "City", "Date.Local", "NO2.Units", "NO2.Mean", "O3.Units", "O3.Mean", "SO2.Units", "SO2.Mean", "CO.Units", "CO.Mean")
cols = c("City", "Date.Local", "NO2.Mean", "O3.Mean", "SO2.Mean", "CO.Mean")
pollution = pollution[cols]
head(pollution)
```
Note that the we have the following units for each pollutant measurement, which we drop from the table for brevity:

* NO2 : parts per billion
* O3 : parts per million
* SO2 : parts per billion
* CO : parts per million

For each city, each day includes four separate rows with differing measurements. We only consider one measurement per day, and drop the rest.
```{r}

```

## Exploratory Data Analysis

## Forecasting

## Conclusion