---
title: "STAT 420 — Project"
author: "Winston Chen, Daniel hooks, Naidan Ganbold, Erik Janc, Michael Lee"
date: "November 19, 2021"
output:
  html_notebook: default
  pdf_document: default
---
```{r setup, include=FALSE}
library(finalfit)
library(knitr)
library(tidyverse)
library(tidyquant)

knitr::opts_chunk$set(echo=TRUE, error=TRUE, eval=TRUE)
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```
## Introduction

## Data Preprocessing
Read the initial CSV file:
```{r}
pollution_raw = read.csv(
  "./data/pollution_us_2000_2016.csv",
  header=TRUE,
)
colnames(pollution_raw)
```
Only include records for the following cities:

* Los Angeles
* New York

Additionally, for ease of viewing, drop irrelevant columns that we are not analyzing.
```{r}
# Subset on relevant cities
cities = c("Los Angeles", "New York")
pollution = subset(
  pollution_raw,
  pollution_raw$City %in% cities,
  drop=TRUE
)

# Subset on relevant columns
cols = c("City", "Date.Local", "NO2.Mean", "O3.Mean", "SO2.Mean", "CO.Mean")
pollution = pollution[cols]
head(pollution)
```
Note that the we have the following units for each pollutant measurement, which we drop from the table for brevity:

* NO2 : parts per billion
* O3 : parts per million
* SO2 : parts per billion
* CO : parts per million

For each city, each day includes four separate rows with differing measurements. We average the columns so that there is only one measurement per day.
```{r}
# Group by city and date, average columns
pollution = pollution %>% 
  group_by(City, Date.Local) %>%
  summarize_each(mean)
pollution$Date.Local = as.Date(pollution$Date.Local, format="%Y-%m-%d")
head(pollution)
```
We expect to see annual seasonality, and in order to appropriately model this, we would require a lag of 365. However, due to a limitation of R that does not allow lag of more than 350 and implementing our own ARIMA method is out of scope of this project. Thus, we instead aggregate on a monthly basis so we can apply a lag of 12 to model annual seasonality.
```{r}
# Aggregate daily data into monthly data
# Save as separate dataframes for each pollutant in each city
aggregate_monthly = function(city, pollutant_idx) {
  df = pollution[pollution$City==city,]
  df$Month = months(df$Date.Local)
  df$Year = format(df$Date.Local, format="%Y")
  if (pollutant_idx == 1) {
    df = aggregate(NO2.Mean ~ Month + Year, df, mean)
    df$Date.Local = as.Date(paste(df$Year, df$Month, "1"), format="%Y %b %d")
    return(xts(df$NO2.Mean, order.by=df$Date.Local))
  } else if (pollutant_idx == 2) {
    df = aggregate(O3.Mean ~ Month + Year, df, mean)
    df$Date.Local = as.Date(paste(df$Year, df$Month, "1"), format="%Y %b %d")
    return(xts(df$O3.Mean, order.by=df$Date.Local))
  } else if (pollutant_idx == 3) {
    df = aggregate(SO2.Mean ~ Month + Year, df, mean)
    df$Date.Local = as.Date(paste(df$Year, df$Month, "1"), format="%Y %b %d")
    return(xts(df$SO2.Mean, order.by=df$Date.Local))
  } else {
    df = aggregate(CO.Mean ~ Month + Year, df, mean)
    df$Date.Local = as.Date(paste(df$Year, df$Month, "1"), format="%Y %b %d")
    return(xts(df$CO.Mean, order.by=df$Date.Local))
  }
}

NY.NO2 = aggregate_monthly("New York", 1)
NY.O3 = aggregate_monthly("New York", 2)
NY.SO2 = aggregate_monthly("New York", 3)
NY.CO = aggregate_monthly("New York", 4)
LA.NO2 = aggregate_monthly("Los Angeles", 1)
LA.O3 = aggregate_monthly("Los Angeles", 2)
LA.SO2 = aggregate_monthly("Los Angeles", 3)
LA.CO = aggregate_monthly("Los Angeles", 4)
```
## Exploratory Data Analysis

```{r}
# Plot time series helpers
plot_single = function(ts, city_idx, pollutant_idx) {
  cities = c("New York", "Los Angeles")
  pollutants = c("NO2", "O3", "SO2", "CO")
  units = c("(parts per billion)", "(parts per million)", "(parts per billion)", "(parts per million)")
  autoplot(
    ts
  ) + labs(
    x="Year",
    y=paste(pollutants[pollutant_idx], units[pollutant_idx]),
    title=paste(cities[city_idx], pollutant[pollutant_idx])
  )
}

plot_together = function(ts1, ts2, pollutant_idx) {
  cities = c("New York", "Los Angeles")
  pollutants = c("NO2", "O3", "SO2", "CO")
  units = c("(parts per billion)", "(parts per million)", "(parts per billion)", "(parts per million)")
  p1 = autoplot(
    ts1
  ) + labs(
    x="Year",
    y=paste(pollutants[pollutant_idx], units[pollutant_idx]),
    title=paste(cities[1], pollutant[pollutant_idx])
  )
  p2 = autoplot(
    ts2
  ) + labs(
    x="Year",
    y=paste(pollutants[pollutant_idx], units[pollutant_idx]),
    title=paste(cities[2], pollutant[pollutant_idx])
  )
  gridExtra::grid.arrange(p1, p2, nrow=2, ncol=1)
}
```

```{r}
plot_together(NY.NO2, LA.NO2, 1)
```

```{r}
plot_together(NY.O3, LA.O3, 2)
```

```{r}
plot_together(NY.SO2, LA.SO2, 3)
```

```{r}
plot_together(NY.CO, LA.CO, 4)
```

```{r}
plot_single(NY.O3, 1, 1)

NY.O3.ddy = diff(diff(log(NY.O3)), 12)[-c(seq(1, 13))]
plot_single(NY.O3.ddy, 1, 1)
```
```{r}
acf(NY.O3.ddy, lag.max=50)
pacf(NY.O3.ddy, lag.max=50)
```
```{r}
# Create various models and compare their AIC
arima(log(NY.O3), order=c(0, 1, 1), seasonal=list(order=c(0, 1, 1), period=12))$aic
arima(log(NY.O3), order=c(0, 1, 2), seasonal=list(order=c(0, 1, 1), period=12))$aic
arima(log(NY.O3), order=c(1, 1, 1), seasonal=list(order=c(0, 1, 1), period=12))$aic
arima(log(NY.O3), order=c(1, 1, 2), seasonal=list(order=c(0, 1, 1), period=12))$aic
arima(log(NY.O3), order=c(1, 1, 1), seasonal=list(order=c(1, 1, 1), period=12))$aic
arima(log(NY.O3), order=c(1, 1, 2), seasonal=list(order=c(1, 1, 1), period=12))$aic
```

```{r}
# Select model and perform model diagnostics
NY.O3.model = arima(log(NY.O3), order=c(1, 1, 1), seasonal=list(order=c(0, 1, 1), period=12))
plot(NY.O3.model$residuals)
acf(NY.O3.model$residuals)
pacf(NY.O3.model$residuals)
qqnorm(NY.O3.model$residuals)
qqline(NY.O3.model$residuals)
```

## Forecasting
```{r}
forecast = predict(NY.O3.model, n.ahead=12)
ts.plot(cbind(ts(NY.O3), exp(forecast$pred)), lty=1:2, main="New York O3 Forecast")
```

## Conclusion